---
title: "PCA with PERMANOVA"
output:
  pdf_document: default
  html_notebook: default
---


## Data import

```{r}
set.seed(27)
data_read<-read.table("Merged_Fold_change_all_values.txt",sep="\t",header=TRUE)
data_read$SampleType<-as.numeric(data_read$SampleType)
data_read$SampleID<-as.numeric(data_read$SampleID)
data_read$Drug<-as.numeric(data_read$Drug)
```

## PCA and PerMANOVA on all data
```{r}
iris_c <- scale(data_read[ ,-1:-4])
library(vegan)
pca <- rda(iris_c)

# plot
plot(pca, type = 'n', display = 'sites', xlim = c(-4,6), ylim=c(-2.5,3.5))
cols <- c('yellow', 'blue')
points(pca, display='sites', col = cols[data_read$SampleType], pch = 16)
ordihull(pca, groups=data_read$SampleType)
ordispider(pca, groups = data_read$SampleType, label = TRUE)

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis(iris_c ~ SampleType, data = data_read, method='eu')

# Coordinates of centroids
a<-ordispider(pca,  groups = data_read$SampleType, label = TRUE)
a[1,]
a[255,]

#Euclidean distance between points
sqrt(((0.1436174 -0.06565365)^2) + ((0.1565998 -0.07158849)^2))

```

The centroid of the patient samples (blue) is (0.1565998,0.1436174). The centroid of the cell line samples (yellow) is (-0.07158849,-0.06565365). The Euclidean ditance between these centroids is 0.1153485.

## PCA and PerMANOVA - outlying cell line and patient samples removed
```{r}
set.seed(27)
data_read<-read.table("Merged_Fold_change_all_values.txt",sep="\t",header=TRUE)
data_read$SampleType<-as.numeric(data_read$SampleType)
data_read$SampleID<-as.numeric(data_read$SampleID)
data_read$Drug<-as.numeric(data_read$Drug)
data_read<-subset(data_read, data_read$SampleID!=32)
data_read<-subset(data_read, data_read$SampleID!=37)
data_read<-subset(data_read, data_read$SampleID!=39)
iris_c <- scale(data_read[ ,-1:-4])
library(vegan)
pca <- rda(iris_c)

# plot
plot(pca, type = 'n', display = 'sites', xlim = c(-4,6), ylim=c(-2.5,3.5))
cols <- c('yellow', 'blue')
points(pca, display='sites', col = cols[data_read$SampleType], pch = 16)
ordihull(pca, groups=data_read$SampleType)
ordispider(pca, groups = data_read$SampleType, label = TRUE)

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis(iris_c ~ SampleType, data = data_read, method='eu')

# Coordinates of centroids
a<-ordispider(pca,  groups = data_read$SampleType, label = TRUE)
a[1,]
a[240,]

#Euclidean distance between points
sqrt(((0.20285899 -0.083530172)^2) + ((0.01003338 -0.004131391)^2))
```
The centroid of the patient samples (blue) is (0.20285899, 0.010033377). The centroid of the cell line samples (yellow) is (-0.08353017, -0.004131391). The Euclidean distance between these points is 0.119474 - removing the outliers moves the centroids slightly further apart.